Covariate analysis of UFL-Mayo-ISB RNASeq (both cerebellum and temporal cortex)

### 2016-05-16 code/Rmd/Mayo_RNASeq.Rmd
Minor sanity fix

```{r diagnosisPlots, fig.height=6, fig.width = 12}
# RIN
p = list()
p[[1]] = ggplot(METADATA, aes(x = Diagnosis, y = RIN, fill = Tissue)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')

@@

Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=10, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E)

@@

WGCNA::plotDendroAndColors(sampleTrees, 
                           labels2colors(cbind(tree,COVARIATES.tmp)),
                           groupLabels = c('Cluster',colnames(COVARIATES.tmp)))
gc()
dev.off()

tmp = paste('Samples in cluster 0 are', paste(sampleTrees$labels[tree == 0], collapse = ','))
tmp = base::gsub('_','\\\\_',tmp)
writeLines(tmp)
tmp = paste('Samples in cluster 3 are', paste(sampleTrees$labels[tree == 3], collapse = ','))
tmp = base::gsub('_','\\\\_',tmp)
writeLines(tmp)
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(VOOM.GENE_EXPRESSION$E, COVARIATES,
                                                 'NULL design(voom-normalized)',
                                                 isKeyPlot=TRUE)

@@

names(clust) = sampleTrees$labels
clust = clust[rownames(COVARIATES)]
```
```{r decompse.normalise.data4, fig.height=10, fig.width=25, results='asis'}
# Tree based clustering
plotDendroAndColors(sampleTrees, 
                    colors = labels2colors(cbind(data.frame(clust = clust), COVARIATES)), 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES))
gc()
dev.off()

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=16}
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.Pathologic_Aging",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.PSP-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.Pathologic_Aging",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.PSP",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.PSP-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.PSP"), 
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

>>>>

```{r diagnosisPlots, fig.height=10, fig.width = 15}
# RIN
p = list()
p[[1]] = ggplot(METADATA, aes(x = Diagnosis, y = RIN, fill = Tissue)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')

@@

Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=15, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E)

@@

WGCNA::plotDendroAndColors(sampleTrees, 
                           labels2colors(cbind(tree,COVARIATES.tmp)),
                           groupLabels = c('Cluster',colnames(COVARIATES.tmp)))
dev.off()

tmp = paste('Samples in cluster 0 are', paste(sampleTrees$labels[tree == 0], collapse = ','))
tmp = base::gsub('_','\\\\_',tmp)
writeLines(tmp)
tmp = paste('Samples in cluster 3 are', paste(sampleTrees$labels[tree == 3], collapse = ','))
tmp = base::gsub('_','\\\\_',tmp)
writeLines(tmp)
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
gc()
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(VOOM.GENE_EXPRESSION$E, COVARIATES,
                                                 'NULL design(voom-normalized)',
                                                 isKeyPlot=TRUE)

@@

names(clust) = sampleTrees$labels
clust = clust[rownames(COVARIATES)]
```
```{r decompse.normalise.data4, fig.height=10, fig.width=25, results='asis'}
# Tree based clustering
COVARIATES.tmp = COVARIATES
COVARIATES.tmp[, FactorCovariates] = sapply(COVARIATES.tmp[,FactorCovariates], as.numeric)
plotDendroAndColors(sampleTrees, 
                    colors = labels2colors(cbind(data.frame(clust = clust), COVARIATES.tmp)), 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES))
gc()
dev.off()

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=25, fig.width=25}
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.Pathologic_Aging",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.PSP-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.Pathologic_Aging",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.PSP",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.PSP-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.PSP"), 
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
